On The Impact of Atoms of Confusion in JavaScript Code

### 2022-07-11 3_Method.tex
The Jupyter Notebook with the analysis of the Repeated Measures Study is finally ready. I have started including the results in the paper.

We call the first study we perform the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et. al~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language. Gopstein et. al. employ a randomized, partial counterbalancing experimental design on a sample of 73 subjects, mostly students, with three or more months of experience in C or C++. The control variable consists of a tiny program containing a single atom candidate and the treatment a functionally-equivalent version that does not contain the atom. The main dependent variable is whether subjects are able to correctly determine the output of the programs. The other dependent variable is the time required to correctly vs. incorrectly determine the outputs. The study has the null hypothesis that \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}. For each atom candidate, six programs were built, three versions including the atom candidate and three versions not including it. Each subject is exposed to four programs for every atom candidate, two including an atom and two not including an atom.

>>>>

\castor{This will become the Repeated Measures Study and we need to introduce the atoms that did not originally appear in Gopstein's study, maybe all of them.}
We call the first study we perform the \textit{``Replication Study''} because it is a partial replication~\cite{Juristo2012} of the first study on atoms of confusion, conducted by Gopstein et. al~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, but targeting JavaScript instead of the C language. Gopstein et. al. employ a randomized, partial counterbalancing experimental design on a sample of 73 subjects, mostly students, with three or more months of experience in C or C++. The control variable consists of a tiny program containing a single atom candidate and the treatment a functionally-equivalent version that does not contain the atom. The main dependent variable is whether subjects are able to correctly determine the output of the programs. The other dependent variable is the time required to correctly vs. incorrectly determine the outputs. The study has the null hypothesis that \textit{``code from both control and treatment groups can be hand-evaluated with equal accuracy''} and the alternate hypothesis that \textit{``the existence of an atom of confusion causes more errors than other code in hand-evaluated outputs.''}. For each atom candidate, six programs were built, three versions including the atom candidate and three versions not including it. Each subject is exposed to four programs for every atom candidate, two including an atom and two not including an atom.

### 2022-07-11 4_Result.tex
The Jupyter Notebook with the analysis of the Repeated Measures Study is finally ready. I have started including the results in the paper.

@@

\subsection{Misunderstanding Rate Analysis}

>>>>

%-----------------------------------------------------------------------
\subsection{Repeated measures study}\label{sec:repeated}

In the repeated measures study, we investigate whether 24 atom candidates and their clear counterparts impact readability. We examine mainly the cases where participants are able to successfully predict the output of a clear version but fail to do so for the corresponding obfuscated version and vice-versa. In addition, similarly to Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, we verify whether there is a trade-off between correctness and time irrespective of the presence of atoms; we check if participants who correctly predict the outputs of the programs tend to do so more slowly than the ones that make incorrect predictions.

Overall, this study had 70 participants. Out of these, 67 have received university-level training in programming, 4 hold a PhD degree, 7 hold a master's degree, and 11 hold a bachelor's degree. 
The median programming experience of the participants is four years (mean 6.5 years) and the median experience with JavaScript programming is one year (mean 2.5 years). 
% General programming: median: 48, 1st quartile: 22,5, 3rd quartile: 91,5
% JavaScript programming: median: 12, 1st quartile: 4, 3rd: 24

%-----------------------------------------------------------------------
\subsection{Correctness Analysis}

For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}. Thus, each participant examined 48 code snippets.  We also measured the time required by the participants to provide their answers. For most cases, participants made more mistakes when attempting to predict the outcomes of the obfuscated versions.

Table~\ref{tab:diff-correctness-repeated} presents the number of correct predictions for the clean and obfuscated versions of each atom candidate, as well as the percentage difference. In only three cases the participants predicted results correctly more often for the obfuscated versions, Arithmetic as Logic, Logic as Control Flow, and Property Access. The former two were empirically validated as atoms of confusion in the study of Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For seven of the atom candidates, the number of correct predictions for the clean versions is more than twice the corresponding number for the obfuscated versions.  Some of these atoms are common programming patterns in JavaScript programs, such as pre- and post-increments and object destructuring. The difference was even greater for Type Conversion (+440\%) and Change of Literal Encoding (+300\%). 
\begin{table}[htbp]
\caption{Summary of the correctness analysis}
\label{tab:diff-correctness-repeated}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
  Atom & Confusing & Clean & $\Delta$(\%)  \\ \midrule
Type Conversion           &            5.0 &      27.0 &                 +440.0 \\
Change of Literal Encoding &           9.0 &      36.0 &                 +300.0 \\
Comma Operator            &            8.0 &      29.0 &                 +262.0 \\
Object Destructuring      &           15.0 &      42.0 &                 +180.0 \\
Assignment As Value       &           18.0 &      43.0 &                 +139.0 \\
Pre-Increment             &           13.0 &      28.0 &                 +115.0 \\
Post-Increment            &            8.0 &      17.0 &                 +112.0 \\
Array Destructuring       &           15.0 &      28.0 &                 +87.0 \\
Indentation no Braces     &           30.0 &      50.0 &                 +67.0 \\
Omitted Curly Braces      &           35.0 &      51.0 &                 +46.0 \\
Object Spread             &           21.0 &      30.0 &                 +43.0 \\
Array Spread              &           22.0 &      31.0 &                 +41.0 \\
Automatic Semicolon Insertion &       15.0 &      21.0 &                 +40.0 \\
Infix Operator Precedence &           47.0 &      53.0 &                 +13.0 \\
Arrow Function            &           54.0 &      58.0 &                  +7.0 \\
Implicit Predicate        &           41.0 &      44.0 &                  +7.0 \\
Indentation With Braces   &           53.0 &      55.0 &                  +4.0 \\
Conditional Operator      &           60.0 &      62.0 &                  +3.0 \\
Constant Variables        &           59.0 &      61.0 &                  +3.0 \\
Dead Unreachable Repeated &           63.0 &      64.0 &                  +2.0 \\
Repurposed Variables      &           23.0 &      23.0 &                   0.0 \\
Arithmetic as Logic       &           52.0 &      47.0 &                 -10.0 \\
Logic as Control Flow     &           33.0 &      28.0 &                 -15.0 \\
Property Access           &           51.0 &      42.0 &                 -18.0 \\
\bottomrule  
\end{tabular}
\end{scriptsize}
}
\end{table}

Data is categorical, paired, dependent. This is analogous to clinical trials where the same patient receives two different treatments. These treatments are spaced to reduce interaction between them. For this, McNemar. Odds ratio as a measure of effect size.

@@

%-----------------------------------------------------------------------
\subsection{Latin-square study}\label{sec:latin}

\subsection{Misunderstanding Rate Analysis}
\castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me which rate is being referred to.}
